---
title: "Using the `lessR` package to investigate wellbeing data"
author: "Stuart Leeds"
date: "01/12/2021"
output: 
    html_document:
        toc: FALSE
bibliography: [../common/packages.bib, ../common/my_library.bib]
csl: ../common/apa.csl
link-citations: true
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}

# Setup
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = "hide",
  fig.align = "center"
)
```

```{r libraries}



# Libraries
library(UsingR) # For wellbeing data
library(lessR) # Main package
library(kableExtra) # For tables
library(lubridate) # For date & time functions

# Creation of R package citation file
 knitr::write_bib(c(.packages(), "rmarkdown"), "../common/packages.bib")
```
<br>

# Introduction

Employee wellbeing is a concern in occupational psychology. Negative
antecedents such as toxic work environments, poor organisational climate,
bullying, amount and type of working hours among many others
[@colliganWorkplaceStressEtiology2006], increase employee stress. The addition
of the Covid-19 pandemic brings new stressors, such as changing working
practices by working from home or being on furlough, which have helped
contribute to what has become known as ["The Great Resignation"
(TGR)](https://www.theguardian.com/money/2021/nov/01/the-great-resignation-almost-one-in-four-workers-planning-job-change).

It would seem that the pandemic has allowed some people to re-evaluate their
lives and working practices, thereby, looking after their own wellbeing by doing
what is best for them. Though these might be troubling times as far as some
employees or organisations are concerned, some recruitment writers see suitable
benefits for how TGR can potentially [benefit your
career](https://www.reed.co.uk/career-advice/the-great-resignation-and-what-it-means-for-your-career/)
and consequently increase wellbeing.

However, this is not an essay or report on wellbeing in the workplace. The
subject is far too deep and complicated to go into here in any great detail, but
the introduction sets the scene.

The idea here is to investigate the relationships between wellbeing and working
hours using the `lessr` package [@lessR2021] which I have recently found to be
very useful.

```{r data}

# Data loading from UsingR
mydata <- wellbeing

# Write to `lessr` .xlsx format (uncomment - only needs doing once)
# Write("./data/wellbeing", format = "Excel", data = mydata)

# Read data
d <- Read("./data/wellbeing.xlsx")
```

# The Data

The wellbeing data are included in the `UsingR` package and originally found
[here](http://prcweb.co.uk/lab/what-makes-us-happy/), where you can see
interactive correlations with wellbeing and variables of your choice. There are
also links to the data origin at [The National Accounts of
Well-being](http://www.nationalaccountsofwellbeing.org/) and category data from
[Gapminder](https://www.gapminder.org/data/), where you can also download other
data from the *Gapminder* database in CSV or XLSX formats.

The wellbeing data consists of $22$ observations (European countries):

> `r d$Country`

and $11$ additional variables:

> `r names(d[-1])`

The data require minimal cleaning, so in this case only the necessary columns
are selected with `d[]` function; and the Well.being column sorted with the
`Sort()` function in `lessR`. For the purposes of using `lessR`, the
`UsingR::wellbeing` data frame was saved as an `.xlsx` file with `Write()` and
re-imported with `Read()` into the variable `d` (you might not have to do this -
I could not get the original data to be manipulated with `lessR` functions).


```{r organising}

# Select columns of interest
d <- d[, .("Country", "Well.being", "Working.hours")]

# Sort countries by wellbeing
d_sort <- Sort(d, Well.being, direction = "-")
```

# The Analysis

Table 1 shows the top five European countries with the highest wellbeing
(`r d_sort$Country[1]`: `r max(d_sort$Well.being)`), and the bottom five with
the lowest wellbeing (`r d_sort$Country[22]`: `r min(d_sort$Well.being)`).
There's not much difference between the highest and lowest positions 
(`r max(d_sort$Well.being)-min(d_sort$Well.being)`).


```{r table1, results='markup'}
# Table top/bottom five
wb_tab <- kbl(list(head(d_sort, 5), tail(d_sort, 5)),
  caption = "<b>Table 1.</b><br><i>European Countries with Highest & Lowest  Wellbeing</i>",
  row.names = F, digits = 2
) |>
  kable_classic(full_width = FALSE) |>
  add_header_above(c(
    "Highest Wellbeing",
    "Lowest Wellbeing"
  ),
  align = "l"
  )

footnote(wb_tab, 
         general = "Working.hours are the average working hours per week per person")
```

\

Not all functions from `lessr` are necessary for this project. So aside from the "tidying" functions used above, the additional functions of interest are
`Regression()` and `regPlot()`.

```{r regression}
# The `Regression` function
r <- Regression(Well.being ~ Working.hours,
  digits_d = 2,
  # Rmd = "lessr_wellbeing_out", # uncomment for Rmd, change name to suit
  graphics = FALSE # Replace with TRUE for three plots:
  # (1: Regression, 2: Distribution, 3: Fitted Values)
)

r
```


The `Regression()` function is really useful. As the name `lessR` suggests, less
$R$ and more output. We want to explore the relationship between wellbeing
(*Intercept*) and the average working hours per week per person (*predictor*).
The output for `Regression()` includes everything you need (and more), all of
which can be presented separately if assigned to its own variable, such as:

\

1. The regression estimate:  
```{r out_estimates, results='markup'}
# The estimates output
r$out_estimates
```
\

2. The Anova:
```{r anova, results='markup'}
# The ANOVA output
r$out_anova
```
\

3. The adjusted R-squared value is $(R_{adj}^2 = `r round(r$Rsqadj, 4)`)$.

\

The statistical information is reported as follows:

> A simple linear regression was carried out to test if Working Hours
> significantly predicted Wellbeing. The results of the regression indicated
> that the model explained `r round(r$Rsqadj*100, 2)`% of the variance and that
> the model was significant, $F$(`r r$anova_model[1]`, `r r$anova_residual[1]`)
> $=$ `r round(r$anova_model[4], 2)`, $p=$ `r r$pvalues[2]`. It was found that
> Working Hours significantly predicted Wellbeing ($B_1=$
> `r round(r$coefficients[2], 2)`, $p=$ `r r$pvalues[2]`).  
> The final predictive model is: proportion of Wellbeing $=$
> `r round(r$coefficients[1], 2)` + (`r round(r$coefficients[2], 2)`
> $\times$ Working Hours)

# Plots

Three plots for visualising the regression and the assumptions are also produced
(automatically in the `Regression()` function, or separately using `regPlot()`), 
as here:

__Figure 1.__  
_Scatterplot showing regression line, prediction and confidence intervals:_

```{r plot1, fig.show='asis'}
# The regression plot
regPlot(r, 1, res_rows = "all") # Max rows = 20, use `res_rows='all'` for all rows
```
\

__Figure 2.__  
_Bar/density plot for distribution of residuals_

```{r plot2, fig.show='asis'}
# The residual barplot
regPlot(r, 2, res_rows = "all") # Max rows = 20, use `res_rows='all'` for all rows
```
\

__Figure 3.__  
_Scatterplot for residuals vs. fitted values_

```{r plot3, fig.show='asis'}
# The residual scatterplot
regPlot(r, 3, res_rows = "all") # Max rows = 20, use `res_rows='all'` for all rows
```

# Additional Output

Additional analyses of interest in the `Regression()` output are the correlation
matrix, which shows a negative moderate relationship between wellbeing and
working hours $(r= -0.51)$:

\

1. `r r$out_title_rel`:

```{r corr, results='markup'}
# The correlation matrix
r$out_cor
```
\

2. `r r$out_title_res` (Top five showing here)

```{r residuals, results='markup'}
# The residuals output
head(r$out_residuals, 10)
```
\

3. `r (r$out_title_pred[1])`(Top 5 again)

```{r prediction, results='markup'}
# The prediction output
head(r$out_predict)
```
\

> Another excellent feature of the `Regression()` function is the `Rmd =
> "filename"` parameter which automatically produces an explanatory and
> exploratory `.Rmd` document of the regression analysis that can be edited; and
> an additional `.html` document, which for this analysis you can read
> [here](./lessr_wellbeing_out.html).

# Further Query

Clearly, working more hours in a week reduces wellbeing. We have already seen
that the Wellbeing Mean for this data set is $`r round(mean(d$Well.being), 2)`$.
This could be considered as *peak* wellbeing, so what are the average working
hours per week per person needed to maintain that level of wellbeing?

The prediction output suggests a level of wellbeing at $5.13$ $(95\%\space CI[4.39, 5.87])$ for $32.95$ (round to $33$) average hours worked at item
$7$:

```{r max hours, results='markup'}
# The prediction output for 5.13 wellbeing
r$out_predict[c(1, 6)]
```
\

This prediction can be confirmed with the regression calculation identified
above:

> Wellbeing = `r round(r$coefficients[1], 2)` + (`r round(r$coefficients[2], 2)`
> $\times$ Working Hours) $\space\therefore\space$
> `r round(r$coefficients[1], 2)` + (`r round(r$coefficients[2], 2)`
> $\times$ 33) $=$
> `r round(sum(r$coefficients[1] + r$coefficients[2] * 33), 2)`

If the preferred working hours to maintain wellbeing at $5.13$ is $33$, then the
question is: would it be more beneficial to work a shorter week with longer
hours per day, or a longer week with fewer hours per day? See Table 2
(Calculations converted to time with `lubridate` [@R-lubridate]).

```{r table 2, results='asis'}
# The data frame for speculative working week
time_table <- data.frame(
  "Days" = c(4, 5),
  "Hours" = c(
      seconds_to_period(33 / 4 * 3600), seconds_to_period(33 / 5 * 3600))
)

# The table for speculative working week
kbl(time_table,
  caption = "<b>Table 2.</b><br><i>Speculative Working Week</i>",
  row.names = FALSE, 
  col.names = c("Days per Week", "Hours per Day")
) |>
  kable_classic(full_width = FALSE)
```
\

Of course there are other options, for example, a three day week at
$11$ hours per day, or a six day week at
$5$ hours $30$ minutes per day, not intending to dismiss those
workers who would do more (or less).

This theory does not account for the four countries that work fewer than $33$
hours per week per person and have a lower than average wellbeing: *Belgium,
France, Germany* and the *United Kingdom* (See the lower-left quadrant of
Figure 1; and Table 3). 

```{r low wb or wh}
# Find countries with wellbeing < 5.13 and working hours < 33
ifelse(d$Well.being < 5.13 & d$Working.hours < 33, d$Country, NA)
```

```{r low wb table, results='asis'}
# Data frame for countries with wellbeing < 5.13 and working hours < 33
less_well <- data.frame(
  Country = c(d$Country[2], d$Country[8], 
              d$Country[9], d$Country[22]),
  Well.being = c(d$Well.being[2], d$Well.being[8], 
                 d$Well.being[9], d$Well.being[22]),
  Well.being.diff = c(d$Well.being[2] - 5.13, d$Well.being[8] - 5.13, 
                      d$Well.being[9] - 5.13, d$Well.being[22] - 5.13),
  Working.hours = c(d$Working.hours[2], d$Working.hours[8], 
                    d$Working.hours[9], d$Working.hours[22]),
  Working.hours.diff = c(d$Working.hours[2] - 33, d$Working.hours[8] - 33, 
                         d$Working.hours[9] - 33, d$Working.hours[22] - 33)
)

# Table for countries with wellbeing < 5.13 and working hours < 33
kbl(less_well,
  caption = "<b>Table 3.</b><br><i>European Countries with Wellbeing < 5.13 and Working Hours < 33</i>",
  row.names = F, digits = 2
) |>
  kable_classic(full_width = FALSE)
```
\

In this group, *`r less_well$Country[2]`* has the largest difference in
wellbeing $(`r less_well$Well.being.diff[2]`)$, a $1/4$ of a point less than
average wellbeing. However, all of the wellbeing differences are *very* small;
and the wellbeing scores for each country are within the  bounds of the 
$95\%\space CI[4.39, 5.87]$ of the wellbeing/working hours prediction. The 
country with the greatest working hours difference is *`r less_well$Country[3]`* 
at $`r round(less_well$Working.hours.diff[3], 2)`$, which is $5$ hours $`r .45*60`$
minutes fewer than average. Further research is required to determine the *whys*
and *wherefores*, which is beyond the scope of this report.

# Conclusion

A very brief introduction to employee wellbeing preceded the purpose of using
various functions from the `lessR` package to explore the relationship between
Well.being and Working.hours in $22$ European countries using the
`UsingR::wellbeing` data set. The `lessR` functions used were `Write()`,
`Read()`, `d[]`, `Sort()`, `Regression()` and `regPlot()`. The key
`Regression()` outputs, *estimate, ANOVA* and *adjusted R-squared* were
presented separately to weave into the text, along with three plots to verify
regression assumptions. Additional output for *correlation, residuals and
influence* and *prediction error* followed. Further query suggested a predicted
number of average working hours to maintain average wellbeing, with speculative
thought on how a working week could be arranged. Finally, the four countries
showing less than wellbeing average and fewer than average working hours were
addressed. In general, `lessR` was found to be an excellent $R$ package to
manipulate data with minimal use of $R$ for maximum output.

# References