---
title: "Final Project for BSc (Honours) Social Psychology: A Summary"
# author: "Stuart Leeds"
date: "14/11/2021"
output:
  html_document:
    toc: no
bibliography: ../common/occ_psych.bib
csl: ../common/apa.csl
link-citations: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center"
)
```

```{r libraries}
library(tidyverse)

```


<br>

<center>**Volunteering and identity: a discursive approach to how  
volunteers see themselves and are seen by others**

<br>

Stuart M. Leeds

<br>

The Open University (UK)</center>

<br>
<br>

```{r data}
interep <- tribble(
  ~`Pos_Neg`, ~`Description`, ~`Perspective`, ~`Regional`, ~`Subject`,
  "Positive", "Privileged positions", "Volunteer", "National (UK)", "Being of importance",
  "Positive", "Privileged positions", "Volunteer", "English", "Being of importance",
  "Positive", "Privileged positions", "Volunteer", "Scottish", "Being of importance",
  "Positive", "Privileged positions", "Volunteer", "English", "Being of importance",
  "Positive", "Privileged positions", "Volunteer", "English", "Being of importance",
  "Positive", "The value of youth", "Volunteer", "Welsh", "Connected importance",
  "Positive", "The value of youth", "Volunteer", "English", "Connected importance",
  "Positive", "The value of youth", "Volunteer", "Scottish", "Connected importance",
  "Positive", "Exalted levels", "Management", "Scottish", "Needy; Worthy; Valuable",
  "Positive", "Exalted levels", "Management", "Scottish", "Needy; Worthy; Valuable",
  "Positive", "Exalted levels", "Management", "English", "Needy; Worthy; Valuable",
  "Positive", "Essential requirement", "Reporter", "English", "Essential; Commited; Indispensible",
  "Positive", "Essential requirement", "Reporter", "Welsh", "Essential; Commited; Indispensible",
  "Positive", "High accolade", "Reporter", "English", "Representative; Belonging; Dependable",
  "Positive", "High accolade", "Reporter", "English", "Representative; Belonging; Dependable",
  "Negative", "Inferior placements", "Volunteer", "National (UK)", "Insignificant; Worthless",
  "Negative", "Inferior placements", "Volunteer", "National (London)", "Insignificant; Worthless",
  "Negative", "Inferior placements", "Volunteer", "Scottish", "Insignificant; Worthless",
  "Negative", "Inferior placements", "Volunteer", "English", "Insignificant; Worthless",
  "Negative", "Negative status", "Management", "National (UK)", "Of little importance",
  "Negative", "Negative status", "Management", "National (London)", "Of little importance",
  "Negative", "Negative status", "Management", "Scottish", "Of little importance",
  "Negative", "Disgruntled people", "Reporter", "National (London)", "Quitters; Unhapy; Volatile",
  "Negative", "Disgruntled people", "Reporter", "English", "Quitters; Unhapy; Volatile"
)
```

**Purpose:** To explore how volunteer identity can be constructed by volunteers
themselves and by others.

**Design:** A Critical Discursive Psychology (CDP) approach using ten local and
national news articles from the Nexis database (word range 5000-7000), found
using the search criteria *‘volunteer’, ‘identity’, ‘interview’* and
*‘charity’*.

Research question:

>_How are volunteer identities constructed in UK news articles?_

The **background** begins with the importance of identity in how we categorise
ourselves and others to help understand who we are. Examples from contemporary
social psychology suggesting the fluidity and complexity of identity are given.
**Previous research** reflects on an earlier literature review assignment which
revealed three themes of identity: *individual, role* and *national*.
**Strengths:** variety of methodologies and methods, quantity and age-ranges of
participants; **limitations:** sample sizes, demographics, disasters between
surveys affecting national identity, selection biases; and **future directions**
of the previous research are discussed.

The **current research** section explains why a CDP approach is used; and
clarifies the meanings of the analytical concepts for identifying construction, 
variation and function in discourse using subject positions made available through interpretive repertoire, and ideological dilemma, to help understand how volunteer identities are constructed in selected news articles. The criteria and process for sourcing the data via the *Nexis* database are explained in the **method** section. The resulting corpus (10 items; 5590 words) included articles in local newspapers from England (4), Scotland (3), Wales (1); and National papers (2). No suitable items were available for Northern Ireland due to content or word count constraints.

**Ethical considerations** were minimal as there were no active participants,
so mainly mentioned awareness of researcher safety and reducing the potential
for harm (psychological distress from potentially reading about volunteer work
in traumatic situations) due to unknown subject-matter. Data searching, storage
and destruction methods are included, along with anonymising any named
individual in articles. For **transparency** the articles are publicly available
for checking credibility of analysis; **reflexivity** involves 'stepping back'
from personal values and volunteering experiences, although some assumptions of
subject knowledge are difficult to avoid.

The **analytical procedure** expands the process of reading/re-reading
the data for familiarisation; and of collating and categorising the data with
*NVivo* to identify and organise the interpretive repertoires, subject
positions and ideological dilemmas.

<br>

**Analysis** (see Figure 1 for a visual summary): It became clear that the news
articles construct volunteer identities from three different perspectives:
*Volunteer; Management*; and *Reporter*:

<u>**Volunteer repertoires.**</u> *Volunteer identity constructed by themselves*

1.  `r interep$Description[1]`
      - Subject positioning: `r interep$Subject[1]`  
        Five extracts see volunteers understanding that their volunteering
        identity is important to themselves, clients, colleagues and
        organisations they volunteer for. The positive subject position is
        constructed from being "head-hunted" for the job, pride in helping the
        community, enjoying the work, team-building, sense of belonging and
        friendship building.

\vspace{0.5cm}

2.  `r interep$Description[17]`
      - Subject positioning: `r interep$Subject[17]`  
        Four extracts show how volunteers have constructed negative identities
        for themselves. These are reflected in "feeling deflated" or "small,
        like an ant"; and being worn out; ignored; or having work dismissed as
        not important.

\vspace{0.5cm}

3.  `r interep$Description[6]`
      - Subject positioning: `r interep$Subject[6]`  
        The three extracts here construct a positive identity for younger
        volunteers who see themselves as approachable and good listeners on an
        equal basis with the youth who use their service.

<u>**Management repertoires.**</u> *Volunteer identity constructed by
management*

1.  `r interep$Description[20]`
      - Subject positioning: `r interep$Subject[20]`  
        A negative identity is constructed here as management position the
        volunteers as, of little importance, unprofessional and unwanted. One
        volunteer feels that management looks down on them as though they were
        the bottom of the hierarchy. Elsewhere, managers find it easier to 'get
        rid' of volunteers rather than train them properly; and that volunteers
        are replaceable. Three of these articles attempt to repair the
        situations.

\vspace{0.5cm}

2.  `r interep$Description[9]`
      - Subject positioning: `r interep$Subject[9]`  
        A positive identity constructed by management as they honour volunteers
        with praise and reward.

<u>**Reporter repertoires.**</u> *Volunteer identity constructed by reporters*

1.  `r interep$Description[23]`
      - Subject positioning: `r interep$Subject[23]`  
        This negative identity is constructed by reporters, though in both
        extracts, management created the circumstances. However, one extract
        relates to a volunteers misunderstanding of a policy, who continues her
        expression of anger regardless.

\vspace{0.5cm}

2.  `r interep$Description[12]`
      - Subject positioning: `r interep$Subject[12]`  
        Reporters have positioned the volunteers with a positive identity by
        writing about one-to-one attention given; and volunteer commitment
        showing dedication and reliability. Positive phraseology and that
        charities struggle to help people without volunteers qualifies the
        volunteers purpose.

\vspace{0.5cm}

3.  `r interep$Description[14]`
      - Subject positioning: `r interep$Subject[14]`  
        More positive subject positioning of volunteer's identity by reporters is
        highlighted in this final interpretive repertoire. Here volunteers are
        exalted and rewarded for their essential work and recognised for their
        length of service.

<br>

```{r plot, out.width = "90%", fig.cap="__Figure 1.__ Interpretive Repertoire Instances for each Perspective by Newspaper Region"}

ggplot(interep, aes(Description, fill = Pos_Neg)) +
  geom_dotplot(
    dotsize = 2.5,
    stackdir = "centerwhole",
    stackgroups = TRUE
  ) +
  facet_grid(c(~Perspective, ~Regional)) +
  coord_flip() +
  theme_bw() +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    legend.position = "bottom"
  ) +
  labs(
    x = "Interpretive Rerpertoire",
    y = "",
    caption = "Source: Author's own",
    fill = "Key"
  ) +
  scale_fill_brewer(direction = -1, palette = "Set2")
```


*Note.* The image was not part of the original document but derived from tables
in the original appendices for this summary. It clearly represents the
interpretive repertoires attributed by each news source (vertical) to each
perspective (horizontal). For example, negative volunteer identities are
reported in three regional areas by management for
"`r interep$Description[20]`"; the Welsh regional paper could be seen as the
most positive with no negative subject positioning; and volunteers were the most
positive *and* negative about themselves.

<br>

**General variation** includes using different phrases to infer age differences.
For example, young, retired, 30 years voluntary service, in their twenties; no
upper limit, are found in eight of the papers. Of particular interest are three
metaphorical uses of the word blood to convey different messages:

1.  "Made my blood boil" (anger).
2.  Calls for "fresh blood" (recruitment drive).
3.  Being the "life-blood" (volunteer importance).

**General function** Several *micro-level* functions exist in the news articles
as contradictions or disagreements. For example, one volunteer positions
themselves as `r interep$Subject[1]` yet feels insignificant as management
'looks down' on them. No *macro-level* (cultural/historical) functions were
evident, although minor cultural differences can be assumed as the articles
cover different regions of England, Scotland and Wales. See figure 1 for the
variation in the spread of positive and negative attribution.



*****

Please use the [email link](mailto:sleeds40@gmail.com) above for questions or
further information.

<br>
