---
title: "Places I have lived"
author: "Stuart Leeds"
date: "2021-07-14"
output: 
  html_document:
    toc: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)
```

>It occurred to me that I have lived in many places during my lifetime. This is a basic run-down of them all, giving rough locations as coordinates. I might return to it at some point and add dates, but this will do for now.

Let's begin by loading the libraries for the code:

```{r libraries}
library(tidyverse)
library(maps)
library(mapproj)
library(leaflet)
library(kableExtra)
library(ggrepel)
```

(_Side-note_. I'm sure there are easier, more efficient ways to do this, I might re-visit and cut down the code). Anyway...

# 1. Locations

These are the places I have lived. Let me state that the list only includes places where I have lived with parents as a child and young adult (items 1:8); and places that I have officially rented or mortgaged as an adult (items 9:end). Furthermore, if memory serves, we lived in two properties in Kuwait, two in Bahrain; and I rented twice in Addlestone.

I have not included sofa-surfing or temporary dwellings. The list represents the order of properties lived in, from birth to now (some 50 years later!)

>_A note on the code to generate the tables and plots in this document: feel free to use it and change details to suit your needs - or improve it (if you do, please email me to say how you did it better - we're all learning, right?)_.

First, the data frame...

```{r dwellings}

# create data frame
dwellings <- data.frame(
  List = 1:18,Address = c(
    "Abridge, Essex",
    "Stansted, Essex",
    "Kuwait City, Kuwait",
    "Manama, Bahrain",
    "Pilgrims Hatch, Essex",
    "Chertsey, Surrey",
    "Stradishall, Suffolk",
    "Cowlinge, Suffolk",
    "Addlestone, Surrey",
    "Twickenham, Middlesex",
    "Twickenham, Middlesex",
    "Bedfont Green, Middlesex",
    "Jaywick, Essex",
    "Harwich, Essex",
    "Harwich, Essex",
    "Ipswich, Suffolk",
    "Bath, Somerset",
    "Bath, Somerset"
  ),
  Latitude = c(
    51.651447,
    51.9070471,
    29.3797091,
    26.2235041,
    51.6357161,
    51.378971,
    52.139740,
    52.158376,
    51.371280,
    51.438541,
    51.434900,
    51.450761,
    51.776301,
    51.933189,
    51.932645,
    52.068666,
    51.394750,
    51.392352
  ),
  Longitude = c(
    0.124044,
    0.2017505,
    47.9735629,
    50.5822436,
    0.2945186,
    -0.520269,
    0.516622,
    0.506907,
    -0.489646,
    -0.336711,
    -0.332810,
    -0.439600,
    1.119079,
    1.274589,
    1.275031,
    1.130408,
    -2.388744,
    -2.384639
  )
)
```

It's technically pointless to include a table because you can see all the variable data above, i just like tables - and you can see the information at a glance, instead of trying to find and link each row in the format above.

```{r main table}
dwellings |>
  kbl(caption = "<b>Table 1</b><br> <i>Places I have lived</i>") |>
  kable_classic_2(full_width = FALSE, position = "left")
```

These next two code-chunks create the variables for the mean and median coordinates, which come in handy later on when we plot them.

```{r mean coords}
meanCoords <- data.frame(
  row.names = c(
    "Worldwide",
    "England"
  ),
  Latitude = c(
    mean(dwellings$Latitude),
    mean(dwellings[-c(3:4), ]$Latitude)
  ),
  Longitude = c(
    mean(dwellings$Longitude),
    mean(dwellings[-c(3:4), ]$Longitude)
  ),
  Place = c(
    "Génicourt-sur-Meuse, France",
    "Cheshunt, Hertfordshire, England"
  )
)
```

```{r median coords}
medianCoords <- data.frame(
  row.names = c(
    "Worldwide",
    "England"
  ),
  Latitude = c(
    median(dwellings$Latitude),
    median(dwellings[-c(3:4), ]$Latitude)
  ),
  Longitude = c(
    median(dwellings$Longitude),
    median(dwellings[-c(3:4), ]$Longitude)
  ),
  Place = c(
    "Upminster, Essex, England",
    "Stapleford Abbotts, Essex, England"
  )
)
```

# 2. Interactive map

The coordinates are pin-pointed on a world map (use +/- zoom functions and left-click[hold]/drag to pan).

* __Orange circles:__ points of all places lived.

* __Green circles:__ _mean_ average location of all coordinates in _Table 1_ (see _Table 2_ for details)

* __Red circles:__ _median_ average locations of all coordinates in _Table 1_ (see _Table 3_ for details)

```{r map}
worldMap <- with(dwellings, leaflet() |>
  addTiles() |>
  addCircleMarkers(
    lng = Longitude,
    lat = Latitude,
    color = "darkorange", label = List
  )) |>
  addCircleMarkers(
    lng = meanCoords[1, 2],
    lat = meanCoords[1, 1], color = "green"
  ) |>
  addCircleMarkers(
    lng = meanCoords[2, 2],
    lat = meanCoords[2, 1], color = "green"
  ) |>
  addCircleMarkers(
    lng = medianCoords[1, 2],
    lat = medianCoords[1, 1], color = "red"
  ) |>
  addCircleMarkers(
    lng = medianCoords[2, 2],
    lat = medianCoords[2, 1], color = "red"
  )

worldMap
```

# 3a. Static plots

## Global view of places lived

```{r map plot}
ggplot(dwellings, aes(x = Longitude, y = Latitude)) +
  geom_point(
    pch = 19,
    colour = dwellings$List,
    size = 2
  ) +
  coord_map(projection = "globular") +
  geom_vline(xintercept = -0.0076589) +
  geom_hline(yintercept = 51.4825766) +
  labs(
    title = "Coordinate plot: Worldwide",
    caption = "Note. Coordinates intercept at Greenwich, London"
  ) +
  theme(plot.caption = element_text(hjust = 0, face = "italic"))
```

## The expanse of locations in England

```{r map uk plot}
ggplot(dwellings, aes(x = Longitude, y = Latitude)) +
  xlim(-2.5, 1.5) +
  ylim(51, 52.5) +
  geom_point(
    colour = dwellings$Latitude,
    size = 3
  ) +
  geom_text_repel(
    label = dwellings$List
  ) +
  coord_map(projection = "globular") +
  geom_vline(xintercept = -0.0076589) +
  geom_hline(yintercept = 51.4825766) +
  labs(
    title = "Coodinate plot: England",
    caption = "Note. Coordinates intercept at Greenwich, London"
  ) +
  theme(plot.caption = element_text(hjust = 0, face = "italic"))
```

You can see that the places lived in England have mainly been between $51^{\circ}$ and $52^{\circ}$ latitude; and between $-2.5^{\circ}$ and $1.5^{\circ}$" longitude (quite a small area, considering). The numbers represent the order in which the places were lived.

It is interesting that although numerous; and despite the huge leap between locations 2 and 5 (3 and 4 abroad), the places lived are in relatively tight clusters suggesting that I've enjoyed a local familiarity for a while before needing/having to move on.

The clusters are divided across counties, as follows. 


# 3b. Breaking down into counties: 

## Essex and Suffolk

```{r map Essex plot}
ggplot(dwellings, aes(x = Longitude, y = Latitude)) +
  xlim(-0.1, 1.3) +
  ylim(51.4, 52.2) +
  geom_point(
    colour = dwellings$List,
    size = 3
  ) +
  geom_text_repel(
    label = dwellings$List
  ) +
  coord_map(projection = "globular") +
  geom_vline(xintercept = -0.0076589) +
  geom_hline(yintercept = 51.4825766) +
  labs(
    title = "Coodinate plot: Essex and Suffolk",
    caption = "Note. Coordinates intercept at Greenwich, London"
  ) +
  theme(plot.caption = element_text(hjust = 0, face = "italic"))
```

## Surrey and Middlesex

```{r map Surrey/middx plot}
ggplot(dwellings, aes(x = Longitude, y = Latitude)) +
  xlim(-0.55, 0.001) +
  ylim(51.35, 51.49) +
  geom_point(
    colour = dwellings$List,
    size = 3
  ) +
  geom_text_repel(
    label = dwellings$List
  ) +
  coord_map(projection = "globular") +
  geom_vline(xintercept = -0.0076589) +
  geom_hline(yintercept = 51.4825766) +
  labs(
    title = "Coodinate plot: Surrey/Middlesex",
    caption = "Note. Coordinates intercept at Greenwich, London"
  ) +
  theme(plot.caption = element_text(hjust = 0, face = "italic"))
```

## Somerset

```{r map Bath plot}
ggplot(dwellings, aes(x = Longitude, y = Latitude)) +
  xlim(-2.4, 0.01) +
  ylim(51.3, 51.5) +
  geom_point(
    colour = dwellings$List,
    size = 3
  ) +
  geom_text_repel(
    label = dwellings$List
  ) +
  coord_map(projection = "globular") +
  geom_vline(xintercept = -0.0076589) +
  geom_hline(yintercept = 51.4825766) +
  labs(
    title = "Coodinate plot: London to Somerset",
    caption = "Note. Coordinates intercept at Greenwich, London"
  ) +
  theme(plot.caption = element_text(hjust = 0, face = "italic"))
```

# 3c. Mean and Median locations

I thought it would be interesting to know the mean and median locations of where I have lived in the world; and in England.

Table 2 shows the mean coordinates of both the worldwide locations; and the England locations from Table 1. In my mind, these give an indication of where I  should be living according to the average coordinates of where I actually _have_ lived.

I quite fancy __Génicourt-sur-Meuse, France__. I've never been there, but it sounds like it would suit me. Ironically, I worked in Cheshunt for a fair while until I was made redundant.

## Mean coordinates

```{r mean coords table}
meanCoords |>
  kbl(caption = "Table 2: Average coordinates (mean)") |>
  kable_classic_2(full_width = FALSE, position = "left")
```

```{r mean avg coords plot}
ggplot(meanCoords, aes(x = Longitude, y = Latitude)) +
  xlim(-1, 10) +
  ylim(49, 53) +
  geom_point(
    colour = "green",
    size = 5
  ) +
  geom_text_repel(
    label = meanCoords$Place,
    nudge_x = 3, nudge_y = 0.5
  ) +
  coord_map(projection = "globular") +
  geom_vline(xintercept = -0.0076589) +
  geom_hline(yintercept = 51.4825766) +
  labs(
    title = "Coodinate plot: the average (mean) of all coordinates",
    caption = "Note. Coordinates intercept at Greenwich, London"
  ) +
  theme(plot.caption = element_text(hjust = 0, face = "italic"))
```

## Median coordinates

Table 3 shows the median coordinates of both the worldwide locations; and the England locations from Table 1. So these will be the absolute mid-point places to live according to both options.

```{r median coords table}
medianCoords |>
  kbl(caption = "Table 3: Average coordinates (median)") |>
  kable_classic_2(full_width = FALSE, position = "left")
```

<br>

>_Note._ The irony of the median coordinate plot is that both locations from the worldwide and England coordinates are a stones throw from where I was born! (Shown with an asterisk on the following map - see note).

```{r med coords plot}
ggplot(medianCoords, aes(x = Longitude, y = Latitude)) +
  xlim(-1, 2.5) +
  ylim(51, 52) +
  geom_point(
    colour = "red",
    size = 5
  ) +
  geom_text_repel(
    label = medianCoords$Place, 
    nudge_x = 1, nudge_y = 0.1
  ) +
  coord_map(projection = "globular") +
  geom_vline(xintercept = -0.0076589) +
  geom_hline(yintercept = 51.4825766) +
  geom_point(x = 0.176, y = 51.5655, pch = 8, size = 2) +
  labs(
    title = "Coodinate plot: the average (median) of all coordinates",
    caption = "Note. Coordinates intercept at Greenwich, London
            * = Rush Green, Dagenham"
  ) +
  theme(plot.caption = element_text(hjust = 0, face = "italic"))
```

So for fun, we'll use the `distm` function from the `geosphere` package to calculate the distances between coordinates in metres (found on [Statistics Globe](https://statisticsglobe.com/geospatial-distance-between-two-points-in-r) - actually a good site for many explanations of R-functions).

Once installed and loaded, I realised I needed another variable to hold the _Rush Green _ coordinates. Then, I found out that `distm` needs the longitude and latitude in that order, so had to reverse the two columns in my original mean/median variables, then combined the two new variables into one.


```{r}
# install.packages("geosphere")

# load additional "geosphere" library
library(geosphere) 

# create rush_green coordinate variable
rush_green <- c(Longitude = 0.176, 
                Latitude = 51.5655, 
                Place = "Dagenham, Essex, England") 

# create new median coords variable and swap long. and lat. variables because that's how `distm` likes it
median_coord_swap <- medianCoords |> 
  relocate(Longitude, .before = Latitude)

#  and combine row 2 with rush_to_stapleford variable
rush_to_stapleford <- median_coord_swap[2,] |> 
  rbind(rush_green)

# calculate distance between median places (Apply distm function)
dist_median <- distm(rush_to_stapleford[, 1:2], fun = distGeo)

# Print distance matrix
dist_median
```
...and the same for the mean of places...

```{r}

# new rush_to_upminster variable combined with row 1 of median swap variable
rush_to_upminster <- median_coord_swap[1,] |> 
  rbind(rush_green)

# calculate mean distance
dist_mean <- distm(rush_to_upminster[, 1:2], fun = distGeo)

# print matrix
dist_mean

```

## Summary

I used [this website](https://www.convertunits.com/from/meters/to/miles) (others are available!) to convert metres to miles (not sure if the site specifies US miles, as others have done, or preferred UK miles (is there a difference?) - but it's good enough for this purpose)

>It's enough to give a guide to how close _Rush Green_ is to each of the median locations:  
* Stapleford Abbotts = 8.73Km (5.43 miles) away  

>and to the median point of the worldwide locations...  
* Upminster = 5.58Km (3.47 miles) away.

This has been an interesting project - thanks for reading. I might take it further one day; and tell the story of each place and why I moved so many times. I'm sure that the average moves during a lifetime is far fewer than what I have achieved here. Does it mean I haven't settled anywhere? Or is it acting on circumstances? A mixture of both? Why have I laid my hat in; and called so many different locations, home? 

(Social) psychology describes these transitional moments of upheaval as `ruptures` [(Zittoun, 2006, p. 5)](https://www.google.co.uk/books/edition/Transitions/ivgnDwAAQBAJ?hl=en&gbpv=1&pg=PR1&printsec=frontcover), since they can cause stress and disruption to 'normal' life (whatever _that_ is). Maybe it explains a few things?